Analyzing Factors Affecting the Potato Prices in Canadian Grocery Markets

### 2024-12-02 paper/paper.qmd
headings for the files in the scrip updated

---
title: "Analyzing the Factors Influencing the Price of Potatoes in Canadian Grocery Markets"
subtitle: "Vendor, Historical Pricing, and Sales Month Affect the Current Potato Prices by Using the Bayesian Modeling Approach"
author: 
  - Fangning zhang
thanks: "Code and data are available at: [https://github.com/FangningZhang81/Canadian-Grocery-Price](https://github.com/FangningZhang81/Canadian-Grocery-Price)"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(ggplot2)
library(here)
library(knitr)
library(dplyr)
library(rstanarm)
library(bayesplot)
library(patchwork)

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
```

@@

The results from the Bayesian model indicate that all three factors—vendor, old pricing, and sales month—significantly influence the current price of potatoes. Old price plays a key role on the current price, with previous prices strongly influencing the current price. The results also show that higher-priced potatoes have greater price fluctuations, while lower-priced potatoes have little difference between the current and old prices. Vendors also influence the current price a lot, with some vendors showing consistently higher or lower prices than others.Seasonal variations, as reflected in the sales month, have a more moderate but still important impact, with potato prices generally increasing during certain months, likely due to heightened demand.

@@

The remainder of this paper is structured as follows. @sec-data provides an overview of the dataset used in the analysis, measurement relating to the dataset, cleaning data process and some details about variables of interest. @sec-model contains the Bayesian regression model used to analyze the factors influencing the current price. @sec-result contains data result visualizations and also the results of the Bayesian regression model. @sec-discussion

@@

# Appendix {.unnumbered}
# Linear model details {#sec-linear-detail}
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

>>>>

---
title: "Analyzing Factors Affecting the Potato Prices in Canadian Grocery Markets"
subtitle: "Vendor Type, Historical Pricing, and Sales Month Play Key Roles in Determining Potato Prices through Bayesian Analysis"
author: 
  - Fangning zhang
thanks: "Code and data are available at: [https://github.com/FangningZhang81/Canadian-Grocery-Price](https://github.com/FangningZhang81/Canadian-Grocery-Price)"
date: today
date-format: long
abstract: "In this study, we analyze the factors influencing potato prices in Canadian grocery markets by employing Bayesian regression modeling. The analysis focuses on how variables such as vendor type, historical pricing, and the sales month affect current pricing trends for yellow and white potatoes. Our findings indicate that vendor type and historical pricing significantly impact potato prices, with distinct variations based on sales month. These insights enhance our understanding of potato pricing dynamics, offering valuable information to consumers, retailers, and policymakers."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(ggplot2)
library(here)
library(knitr)
library(dplyr)
library(rstanarm)
library(bayesplot)
library(patchwork)
library(arrow)

analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
```

@@

The results from the Bayesian model indicate that all three factors—vendor, old pricing, and sales month—significantly influence the current price of potatoes. Old price plays a key role on the current price, with previous prices strongly influencing the current price. The results also show that higher-priced potatoes have greater price fluctuations, while lower-priced potatoes have little difference between the current and old prices. Vendors influence the current price a lot, with some vendors showing consistently higher or lower prices than others. Seasonal variations, as reflected in the sales month, have a more moderate but still important impact, with potato prices generally increasing during certain months.

@@

The remainder of this paper is structured as follows. @sec-data provides an overview of the dataset used in the analysis, measurement relating to the dataset, cleaning data process and some details about variables of interest. @sec-model contains the Bayesian regression model used to analyze the factors influencing the current price. @sec-result contains data result visualizations and also the results of the Bayesian regression model. @sec-discussion discusses what was done, implications, limitations and next steps. @sec-appendix includes more details for the models and also an appendix where we focused on some aspect of surveys, sampling and observational data.

@@

# Appendix {.unnumbered} {#sec-appendix}
# Linear model details {#sec-linear-detail}
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
